---
title: <span style="color:steelblue">Predicting Electricity Consumption</span>
author: "Nolan Cardozo"
date: "15 February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###<span style="color:steelblue">Loading the Factbook dataset into R.</span>
```{r}
set.seed(500)
factbook <- read.csv("C:/Users/Admin/Desktop/Machine Learning/ML/factbook.csv", sep=";", stringsAsFactors=FALSE)
``` 


###<span style="color:steelblue">Filling the missing values of the dataset by NA which will help in further imputation.</span>
```{r}
factbook[factbook==""] <- NA
```

###<span style="color:steelblue">Excluding the first row of the dataset which contains the column data type.</span>
```{r}
factbook <- factbook[2:nrow(factbook),]
```


###<span style="color:steelblue">Converting all variables other than country variable to type numeric.</span>
```{r}
for(i in 2:ncol(factbook)){
  factbook[,i]<-as.numeric(factbook[,i])
}
```


###<span style="color:steelblue">Assesing and visualizing missing values using VIM package.</span>

```{r, include=FALSE}
library(VIM)
library(mice)
```

```{r}
imputation_plot<-aggr(factbook,col=c('yellow','red'),
                      numbers=TRUE,sortVars=TRUE,
                      labels=names(factbook),gap=3,
                      ylab=c("Missing Data","Pattern"))
```

###<span style="color:steelblue">Mice function using cart method for imputation.</span>
```{r, include=FALSE}
tempdata <- mice(factbook,m=5,maxit=10,meth='cart',seed =100)
```
tempdata <- mice(factbook,m=5,maxit=10,meth='cart',seed =100)

###<span style="color:steelblue">Extacting one of the 5 imputed sets.</span>
```{r}
completeData<-complete(tempdata,4)
```

###<span style="color:steelblue">Assessing the dataset after imputation.</span>
```{r}
completeData$Electricity...production.kWh. <- NULL
imputation_plot<-aggr(completeData,col=c('yellow','blue'),
                      numbers=TRUE,sortVars=TRUE,
                      labels=names(completeData),gap=1,
                      ylab=c("Missing Data","Pattern"), plot=TRUE)
```

###<span style="color:steelblue">80-20 training - test split for multiple linear regression.</span>
```{r}
set.seed(400)
trainno <- sample(1:nrow(completeData),0.8*nrow(completeData))
train <- completeData[trainno,]
test <- completeData[-trainno,]
```

###<span style="color:steelblue">Multiple linear regression model 1.</span>
```{r}
lm <- lm(Electricity...consumption.kWh. ~ Area.sq.km. + Birth.rate.births.1000.population. + Death.rate.deaths.1000.population. + Infant.mortality.rate.deaths.1000.live.births.,data=train)
summary(lm)
pred1 <- predict(lm,test)
layout(matrix(c(1,2,3,4),2,2))
plot(lm)
```

###<span style="color:skyblue">Validating using RMSE for the model 1.</span>
```{r, include=FALSE}
library(hydroGOF)
```

```{r }
rmse(predict(lm,test),test$Electricity...consumption.kWh.)
```

###<span style="color:steelblue">70-30 training - test split for multiple linear regression.</span>
```{r}
set.seed(400)
trainno2 <- sample(1:nrow(completeData),0.7*nrow(completeData))
train2 <- completeData[trainno2,]
test2 <- completeData[-trainno2,]
```

###<span style="color:steelblue">Multiple linear regression model 2.</span>
```{r}
lm2 <- lm(Electricity...consumption.kWh. ~ Area.sq.km. + Birth.rate.births.1000.population. + Death.rate.deaths.1000.population. + Infant.mortality.rate.deaths.1000.live.births.,data=train2)
summary(lm2)
pred2 <-predict(lm2,test2)
layout(matrix(c(1,2,3,4),2,2))
plot(lm2)
```

###<span style="color:skyblue">Validating using RMSE for the model 2.</span>

```{r }
rmse(predict(lm2,test2),test2$Electricity...consumption.kWh.)
```

###<span style="color:steelblue">Cross-Validation model using DAAG pacage and cv.lm() function.</span>
```{r}
library(DAAG)
```

```{r}
layout(1)
lm3 <-cv.lm(completeData,form.lm=formula(Electricity...consumption.kWh. ~ Area.sq.km. + Birth.rate.births.1000.population. + Death.rate.deaths.1000.population. + Infant.mortality.rate.deaths.1000.live.births.),m=11)
```

###<span style="color:steelblue">Testset using first 20 rows of data and testing it on all the models created.</span>
```{r}
testset <- completeData[1:20,]
testset$model1output<-predict(lm,testset)
testset$model2output<-predict(lm2,testset)
```

###<span style="color:steelblue">Predicted values of 1st 20 values of cross validation model.</span>
```{r}
testset$Cvmodeloutput <- lm3[1:20,]$cvpred
```

###<span style="color:steelblue">Evaluating the 3 models with the actuals.</span>
```{r}
evaloutput <- testset[,c("Electricity...consumption.kWh.","model1output","model2output","Cvmodeloutput")]
evaloutput
```

###<span style="color:steelblue">Evaluating the output it is found that cross validation model performs much better than the 80-20 training test split and 70-30 training test split.</span>
